---
title: "Download Data"
subtitle: "SKILLS: Data Preparation and Workflow Management - Group 9"
author:
  - "Rabino Tommaso"
  - "Franceschini Emanuele"
  - "Magalotti Bianca"
  - "Tan Colin"
  - "Benmrit Akram"
date: "\\textit{20 October 2024}"
output:
  pdf_document:
    latex_engine: xelatex
    toc: true
    toc_depth: 2
    number_sections: true
    fig_caption: true
header-includes:
  - "\\usepackage[english]{babel}"  # Language
  - "\\usepackage[T1]{fontenc}"  # Font encoding
  - "\\usepackage{mathptmx}"  # Times New Roman font for text
  - "\\usepackage{helvet}"  # Arial-like font for sans-serif
  - "\\usepackage{setspace}"  # Line spacing
  - "\\onehalfspacing"  # 1.5 line spacing
  - "\\usepackage{fancyhdr}"  # Header and footer customization
  - "\\usepackage{titlesec}"  # Section titles formatting
  - "\\usepackage{abstract}"  # Abstract formatting
  - "\\usepackage{caption}"  # Captions customization
  - "\\usepackage{graphicx}"  # Graphics
  - "\\usepackage{amsmath}"  # Math equations
  - "\\usepackage{amssymb}"  # Math symbols
  - "\\usepackage{natbib}"  # Citation style (change as needed)
  - "\\bibliographystyle{apalike}"  # Bibliography style (change as needed)
  - "\\usepackage{hyperref}"  # Hyperlinks and URLs
  - "\\hypersetup{colorlinks=true,linkcolor=blue,citecolor=blue,urlcolor=blue}"
  - "\\usepackage{appendix}"  # Appendix formatting
  - "\\usepackage{enumerate}"  # Enumerate environment
  - "\\pagestyle{fancy}"  # Custom page style
  - "\\fancyhf{}"
  - "\\renewcommand{\\headrulewidth}{0pt}"
  - "\\renewcommand{\\footrulewidth}{0pt}"
  - "\\fancyhead[R]{\\thepage}"  # Page number in the header
  - "\\fancypagestyle{plain}{\\fancyhf{}\\renewcommand{\\headrulewidth}{0pt}}"
  - "\\lhead{\\small{A.A. 2023/2024-Courses: SKILLS: Data Preparation and Workflow Management}}"
  - "\\usepackage{multicol}"  # For two columns
geometry: "left=2.5cm, right=1.5cm, top=2.5cm, bottom=2.5cm"  # Adjust margins
---

```{r, include=FALSE}
#Load Required Packages
rm(list=ls())
library(readr)
```

\section{Introduction}

\textbf{Airbnb.} Airbnb is a prominent online marketplace founded in 2008, enabling individuals to lease their homes, apartments, or individual rooms to travelers. Over the years, Airbnb has evolved into a global platform operating in more than 220 countries and regions. Its mission revolves around fostering a world where everyone can find their place anywhere, facilitating unique travel experiences, and offering affordable accommodation options. Airbnb's impact on the hospitality industry has been transformative, providing an alternative to conventional hotels. With a vast global presence boasting over 7 million listings, Airbnb is the preferred choice for travelers seeking distinctive and cost-effective lodgings. The platform caters to a diverse array of accommodation choices, ranging from shared rooms to entire residences, accommodating various preferences such as location, pricing, and amenities. Additionally, Airbnb features a secure messaging system for direct communication between hosts and guests, along with a reliable payment system to facilitate transactions. Consequently, Airbnb has garnered favor among both leisure and business travelers, as well as homeowners seeking to monetize their unused space.

\textbf{Starting Point - The Dataset.} This project centers around the Milan listings dataset, accessible on the official \href{http://insideairbnb.com/get-the-data/}{Inside Airbnb} website. Inside Airbnb serves as Airbnb's official platform for collecting and storing all listing data. The dataset encompasses information on more than 23,000 apartments available for rent in Milan through Airbnb. The following code block provides the download link for this dataset.


\section{Preliminary settings}
Before proceeding with the dataset download and extraction, we need to complete some preliminary steps:

\begin{itemize}
  \item The \texttt{url} object is a vector containing URL pointing to ZIP file hosted by Inside Airbnb. The URL corresponds to Airbnb listing data for Milan.
  \item The \texttt{name} object is another vector providing user-defined name for the URL. These name will be used to label the downloaded data frame, making it easier to work with.
  \item The \texttt{data\_dir} object specifies the target directory where the ZIP file will be downloaded and extracted.
\end{itemize}

These preliminary settings ensure that we have organized access to the dataset we need for our research. With this foundation, we can proceed to acquire, explore and analyze the data.

```{r}
# URL for downloading the Milan dataset ZIP file
url <- "http://data.insideairbnb.com/italy/lombardy/milan/2023-06-21/data/listings.csv.gz"

# Name for the Milan dataset
name <- "listings_milan"

# Define the target directory for extraction
data_dir <- "../../data/"
```


\section{Download and Extraction}
The following code block is responsible for downloading, extracting, and loading the Milan Airbnb listings dataset. It checks if the dataset file already exists locally and, if not, proceeds to download it. After successful download, the code extracts the data from the compressed file and loads it into the 'data_milan' variable. A confirmation message will be printed to indicate that the dataset has been loaded.

```{r}
#Avoid too long downloads to be shut down
options(timeout = max(1000, getOption("timeout")))

# Define the target file path
target_file <- paste0(data_dir, "milan.csv.gz")

# Check if the CSV file already exists
if (!file.exists(target_file)) {
  download.file(url, target_file, mode = "wb")
} else {
  print("ZIP file milan.csv.gz already exists. Skipping download.")
}

# Extract and load the CSV file
gz_file <- gzfile(target_file, "rt")
data_milan <- read.csv(gz_file)
close(gz_file)

# Print a message to confirm loading
print("Loaded milan.csv.gz")
```

\section{Delete Source ZIP Files}
In this section, we address the cleanup process by removing the source ZIP file that was downloaded earlier. This ZIP file contained Milan Airbnb listings data. Once the relevant data have been extracted and the dataset saved locally, theis ZIP file is no longer needed and can be safely deleted to free up storage space. This step ensures that we maintain an organized and clutter-free working environment.
```{r}
# Delete the source ZIP file if it exists
if (file.exists(target_file)) {
  file.remove(target_file)
  print("Deleted ZIP file")
} else {
  print("ZIP file does not exist. Skipping deletion.")
}
```



\section{Save the dataset}
Finally, the code ensures the lasting accessibility of the dataset by saving it in two distinct formats. It is preserved both as a CSV file, facilitating external examination outside the RStudio environment, and as a .RDS (R Data Serialization) object. The latter format is chosen to simplify the reloading process, ensuring seamless integration into subsequent project sections.
```{r}
# Save the data as a CSV file
write.csv(data_milan, file = "../../data/dataset1/dataset.csv", fileEncoding = "UTF-8", row.names = FALSE)

# Save the data as .Rds file
saveRDS(data_milan, file = "../../gen/data-preparation/input/raw_data.rds")
```


```{r, include=FALSE}
rm(list = ls())
```
